---
output:
  pdf_document: default
  html_document: default
---

\section{Literature Review}

\subsection{ARIMAX-GARCHX with Simulated Peaks}

\subsubsection{Autoregressive Moving Average with Exogenous Variables (ARMAX)}
In Adli's research about forecasting Steel Prices, the ARIMAX model seems particularly advantageous as it incorporates both past values of the dependent variable and external factors that may influence these prices (Adli 2023). Moreover, the ARIMAX model has been shown to outperform simpler models in various
applications, including agricultural contexts(Sudipa et al., 2024).

\subsubsection{Generalized Autoregressive Conditional Heteroscedasticity with Exogenous Variable (GARCHX)}
GARCHX models further extend GARCH by including exogenous regressors in the variance equation. This enhancement allows the conditional variance to respond to external factors. Ratih et al. (2018) showed that incorporating macroeconomic indicators in GARCHX improved VaR estimation for stock indices, indicating its usefulness in contexts with external volatility drivers, such as agricultural commodities.

\subsection{Commodity with Price Peaks Modeling Approaches}
Leskinen and Kangas (1998) developed a simulation-based model for timber price forecasting that separates normal dynamics from rare price peaks. The model applied an AR(1) process to the regular component and used simulated shock processes to reflect rare, high-impact events. This dual-component strategy allows for more realistic long-term planning in forest economics, and their peak modeling approach informs the simulation method used in our study.

\subsection{Differentiation of This Method}
This study introduces an integrated ARIMAX-GARCHX framework, enhanced with a simulated peak shock mechanism, to better capture cocoa price dynamics. Our approach differs from standard models in several ways:

\begin{itemize}
  \item It incorporates exogenous weather variables in both the mean and variance equations, capturing direct environmental influences.
  \item Z-score filtering is applied to remove historical extreme price spikes from training data, enhancing robustness.
  \item A lognormal shock process is simulated and superimposed on the forecast horizon to mimic possible future disruptions.
  \item This hybrid structure enables the model to explain both regular market movements and rare, high-impact anomalies, improving forecast realism and flexibility.
\end{itemize}



\section{Methodology}
```{r, echo = FALSE, message = FALSE, warning = FALSE, include = FALSE}
### Load Required Libraries
library(tidyverse)
library(lubridate)
library(forecast)
library(tseries)
library(ggplot2)
library(rugarch)
library(zoo)
library(knitr)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE, include = FALSE}
### Load and Merge Data
cocoa_prices <- read.csv("Daily Prices_ICCO.csv", stringsAsFactors = FALSE)
cocoa_prices$Date <- as.Date(cocoa_prices$Date, format='%d/%m/%Y')
cocoa_prices$Price <- as.numeric(gsub(",", "", cocoa_prices$ICCO.daily.price..US..tonne.))
cocoa_prices <- cocoa_prices %>% select(Date, Price) %>% arrange(Date)

ghana_weather <- read.csv("Ghana_data.csv", stringsAsFactors = FALSE)
ghana_weather$DATE <- as.Date(ghana_weather$DATE)
ghana_weather <- ghana_weather %>%
  group_by(DATE) %>%
  summarise(across(c(PRCP, TAVG, TMAX, TMIN), mean, na.rm = TRUE))

cocoa_data <- left_join(cocoa_prices, ghana_weather, by = c("Date" = "DATE")) %>%
  mutate(log_price = log(Price),
         diff_log_price = c(NA, diff(log_price))) %>%
  drop_na()

# check stationary
library(tseries)
adf.test(cocoa_data$log_price)      
kpss.test(cocoa_data$log_price)     

adf.test(cocoa_data$diff_log_price)
kpss.test(cocoa_data$diff_log_price)

weather_vars <- cocoa_data %>% select(PRCP, TAVG, TMAX, TMIN)

lapply(weather_vars, adf.test)
lapply(weather_vars, kpss.test)

cocoa_data <- cocoa_data %>%
  mutate(
    d_PRCP = c(NA, diff(PRCP)),
    d_TAVG = c(NA, diff(TAVG)),
    d_TMIN = c(NA, diff(TMIN)),
    d_TMAX = TMAX  
  ) %>% drop_na()

# Prepare external regressors matrix (lag to align with diff_log_price)
external_vars <- cocoa_data %>% select(d_PRCP, d_TAVG, d_TMAX, d_TMIN)
external_matrix <- as.matrix(external_vars)

```


```{r, echo = FALSE, message = FALSE, warning = FALSE}
### Identify Peaks via Z-score
threshold <- 3
cocoa_data <- cocoa_data %>%
  mutate(z_score = scale(diff_log_price),
         is_peak = abs(z_score) > threshold)

# Remove peaks and lag effects (5-day window)
lag_window <- 5
cocoa_data$peak_or_lag <- FALSE
peak_indices <- which(cocoa_data$is_peak)
for (i in peak_indices) {
  cocoa_data$peak_or_lag[i:min(i + lag_window, nrow(cocoa_data))] <- TRUE
}
cocoa_data$clean_log_diff <- ifelse(cocoa_data$peak_or_lag, NA, cocoa_data$diff_log_price)
cocoa_data$clean_log_diff <- zoo::na.approx(cocoa_data$clean_log_diff, na.rm = FALSE)

### Train-Test Split
train_size <- floor(0.8 * nrow(cocoa_data))
train_data <- cocoa_data[1:train_size, ]
test_data <- cocoa_data[(train_size + 1):nrow(cocoa_data), ]
log_returns <- cocoa_data$diff_log_price
train_returns <- log_returns[1:train_size]
test_returns <- log_returns[(train_size + 1):length(log_returns)]

# Split external regressors
train_external <- external_matrix[1:train_size, ]
test_external <- external_matrix[(train_size + 1):nrow(external_matrix), ]
test_dates <- cocoa_data$Date[(train_size + 1):length(log_returns)]
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}

### ARIMAX + GARCH-X Model
garchx_spec <- ugarchspec(
  variance.model = list(
    model = "sGARCH", 
    garchOrder = c(1, 1),
    external.regressors = train_external  # GARCH-X 关键点
  ),
  mean.model = list(
    armaOrder = c(1, 1), 
    include.mean = TRUE,
    external.regressors = train_external  # ARIMAX
  ),
  distribution.model = "std"
)

# Fit the model
garchx_fit <- ugarchfit(spec = garchx_spec, data = train_returns)

# Forecast using external regressors for test set
garchx_forecast <- ugarchforecast(
  garchx_fit,
  n.ahead = length(test_returns),
  external.forecasts = list(
    mregfor = test_external,
    vregfor = test_external
  )
)

# Extract predicted returns
predicted_returns <- as.numeric(fitted(garchx_forecast))
predicted_sigma <- as.numeric(sigma(garchx_forecast))

# Reconstruct price series from last training price
last_log_price <- tail(train_data$log_price, 1)
forecast_log_price <- cumsum(predicted_returns) + last_log_price
forecast_price <- exp(forecast_log_price)

# Build forecast dataframe
garchx_df <- tibble(
  Date = test_dates,
  Price = forecast_price
)

# Construct upper/lower bands (±1 conditional sigma)
upper_band <- exp(forecast_log_price + predicted_sigma)
lower_band <- exp(forecast_log_price - predicted_sigma)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# Simulate Peak Shocks using lognormal distribution
set.seed(123)
shock_dates <- sample(1:nrow(test_data), 10)
shock_effect <- rep(0, nrow(test_data))
gamma <- 0.8

for (d in shock_dates) {
  for (h in 0:5) {
    idx <- d + h
    if (idx <= length(shock_effect)) {
      shock_effect[idx] <- shock_effect[idx] + gamma^h * rlnorm(1, meanlog = -3.7, sdlog = 1)
    }
  }
}

forecast_price_shocked <- exp(cumsum(predicted_returns + shock_effect) + last_log_price)

forecast_df <- tibble(
  Date = test_data$Date,
  Price = forecast_price,
  Price_shocked = forecast_price_shocked,
  Upper = upper_band,
  Lower = lower_band
)
```



\subsection*{ARIMAX-GARCHX Model with Simulated Peak Shocks}

\subsubsection*{1. Model Specification}

The ARIMAX-GARCHX model jointly captures the conditional mean and conditional variance of time series using both internal dynamics and external variables. The formulation is as follows:

\[
\begin{aligned}
Y_t &= \mu + \sum_{i=1}^{p} \phi_i Y_{t-i} + \sum_{j=1}^{q} \theta_j \epsilon_{t-j} + \sum_{k=1}^{m} \beta_k X_{k,t} + \epsilon_t \\\\
\epsilon_t &= \sigma_t z_t,\quad z_t \sim \text{i.i.d. } (0,1) \\\\
\sigma_t^2 &= \omega + \alpha \epsilon_{t-1}^2 + \beta \sigma_{t-1}^2 + \sum_{k=1}^{m} \gamma_k X_{k,t}
\end{aligned}
\]

Where:
- \( Y_t \): the differenced log cocoa price \( \log(P_t) - \log(P_{t-1}) \)
- \( \epsilon_t \): the residual from the ARIMAX mean equation
- \( \sigma_t^2 \): the conditional variance
- \( X_{k,t} \): external covariates (weather variables)
- \( \phi_i, \theta_j \): AR and MA coefficients
- \( \beta_k, \gamma_k \): coefficients of exogenous variables in the mean and variance equations, respectively

\subsubsection*{2. Model Justification}

\begin{itemize}
    \item \textbf{Autoregressive Dynamics:} Cocoa price returns exhibit autocorrelation and mean-reverting behavior, warranting an ARMA structure in the mean equation.
    
    \item \textbf{External Covariates:} Weather conditions directly impact cocoa production. We include daily precipitation (PRCP), average temperature (TAVG), maximum temperature (TMAX), and minimum temperature (TMIN) as external regressors.
    
    \item \textbf{Volatility Clustering:} The GARCH(1,1) structure captures observed volatility clustering in returns. Extending it to GARCHX allows weather variability to influence volatility directly.
    
    \item \textbf{Rare Event Simulation:} To test the model’s robustness under extreme conditions, we simulate rare peak shocks during the forecast period using a lognormal decay process.
\end{itemize}


\subsubsection*{3. Simulated Peak Shocks}

To simulate rare exogenous disruptions (e.g., supply crisis), we inject synthetic peak shocks into the test forecast:

\[
s_{t+h} = \gamma^h \cdot \xi_t,\quad \xi_t \sim \text{LogNormal}(\mu_s, \sigma_s^2),\quad h = 0,1,\dots,H
\]

These shocks decay over time and are added to the predicted returns:

\[
\tilde{r}_t = \hat{r}_t + s_t
\quad \Rightarrow \quad
\tilde{P}_t = \exp\left( \log(P_0) + \sum_{i=1}^{t} \tilde{r}_i \right)
\]

\subsubsection*{4. Forecast Visualization}

The final plot displays:
\begin{itemize}
    \item Historical cocoa prices in black
    \item ARIMAX-GARCHX forecast in blue
    \item Shock-adjusted forecast in dashed red
    \item 1-sigma confidence bands as a light blue ribbon
\end{itemize}

A vertical dashed line denotes the beginning of the test period. This visualization demonstrates the model's baseline predictive accuracy and behavior under extreme simulated conditions. These preprocessing steps ensure that the modeling assumptions of ARIMAX and GARCHX are met, particularly with regard to stationarity and the absence of structural breaks or missingness that could bias parameter estimates.

```{r, echo = FALSE, message = FALSE, warning = FALSE}

# Extract actual and predicted values
actual <- test_data$Price
predicted <- forecast_price[1:length(actual)]

# Compute residuals
residuals <- predicted - actual

# === Core Metrics ===
ME   <- mean(residuals)
RMSE <- sqrt(mean(residuals^2))
MAE  <- mean(abs(residuals))
MPE  <- mean(residuals / actual) * 100
MAPE <- mean(abs(residuals / actual)) * 100

# === MASE (using naive one-lag MAE) ===
naive_errors <- abs(diff(train_data$Price))
naive_mae <- mean(naive_errors)
MASE <- MAE / naive_mae

# === Residual autocorrelation ===
ACF1 <- acf(residuals, lag.max = 1, plot = FALSE)$acf[2]

# Compile into summary table
performance_metrics <- tibble(
  Metric = c("Mean Error (ME)", 
             "Root Mean Squared Error (RMSE)", 
             "Mean Absolute Error (MAE)", 
             "Mean Percentage Error (MPE)", 
             "Mean Absolute Percentage Error (MAPE)", 
             "Mean Absolute Scaled Error (MASE)", 
             "Lag-1 Autocorrelation of Residuals (ACF1)"),
  Value = round(c(ME, RMSE, MAE, MPE, MAPE, MASE, ACF1), 4)
)

# Display as clean table
kable(performance_metrics, caption = "Model Performance Metrics on Test Set")
ft <- flextable::qflextable(performance_metrics)
flextable::save_as_image(ft, path = "table/metrics_table.png")
```


```{r, echo = FALSE, message = FALSE, warning = FALSE}
### Plot
forecast_plot <- ggplot() +
  geom_line(data = cocoa_data, aes(x = Date, y = Price), color = "black") +
  geom_line(data = garchx_df, aes(x = Date, y = Price), color = "blue") +
  geom_line(data = forecast_df, aes(x = Date, y = Price_shocked), color = "red", linetype = "dashed") +
  geom_ribbon(data = forecast_df, aes(x = Date, ymin = Lower, ymax = Upper), fill = "blue", alpha = 0.2) +
  labs(
    title = "ARIMAX GARCH-X Forecast with Simulated Price Peaks vs Actual Prices",
    y = "Price", x = "Date"
  ) +
  theme_minimal()+
  geom_vline(xintercept = as.numeric(test_data$Date[1]), linetype = "dashed", color = "gray")

forecast_plot

```


\section{Data}

\subsubsection*{1. Dataset Description}

\subsection*{1. Daily Cocoa Price Data (ICCO)}

This dataset contains the daily prices of cocoa, as published by the International Cocoa Organization (ICCO). It includes:

\begin{itemize}
    \item \textbf{Date:} The date of observation, ranging from the earliest to most recent available day.
    \item \textbf{Price:} The ICCO daily price in US dollars per tonne.
\end{itemize}

The dataset comprises 7,812 records with 7,685 unique price entries. There are minor instances of repeated dates and prices.

\subsection*{2. Ghana Weather Data}

This dataset contains daily weather observations from multiple stations in Ghana. It includes:

\begin{itemize}
    \item \textbf{STATION:} The ID of the observation station
    \item \textbf{NAME:} The name of the station
    \item \textbf{DATE:} The date of the observation
    \item \textbf{PRCP:} Daily precipitation in millimeters. If blank, it indicates no precipitation.
    \item \textbf{TAVG:} Daily average temperature (2 meters above ground)
    \item \textbf{TMAX:} Maximum daily temperature
    \item \textbf{TMIN:} Minimum daily temperature
\end{itemize}

The dataset includes 53,231 records across 10 stations.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# -- Cocoa Price Summary --
cocoa_price_summary <- cocoa_prices %>%
  summarise(
    Count = n(),
    Start_Date = min(Date, na.rm = TRUE),
    End_Date = max(Date, na.rm = TRUE),
    Mean_Price = mean(Price, na.rm = TRUE),
    SD_Price = sd(Price, na.rm = TRUE),
    Min_Price = min(Price, na.rm = TRUE),
    Max_Price = max(Price, na.rm = TRUE),
    Median_Price = median(Price, na.rm = TRUE)
  )

kable(cocoa_price_summary, caption = "Summary Statistics of Cocoa Price Data")
ft <- flextable::qflextable(cocoa_price_summary)
flextable::save_as_image(ft, path = "table/cocoa_price_summary_table.png")



# --️ Ghana Weather Summary --
weather_numeric_summary <- ghana_weather %>%
  select(PRCP, TAVG, TMAX, TMIN) %>%
  summarise(
    Count = n(),
    PRCP_Mean = mean(PRCP, na.rm = TRUE),
    PRCP_SD = sd(PRCP, na.rm = TRUE),
    TAVG_Mean = mean(TAVG, na.rm = TRUE),
    TAVG_SD = sd(TAVG, na.rm = TRUE),
    TMAX_Mean = mean(TMAX, na.rm = TRUE),
    TMAX_SD = sd(TMAX, na.rm = TRUE),
    TMIN_Mean = mean(TMIN, na.rm = TRUE),
    TMIN_SD = sd(TMIN, na.rm = TRUE)
  )

kable(weather_numeric_summary, caption = "Summary Statistics of Ghana Weather Variables")
ft <- flextable::qflextable(weather_numeric_summary)
flextable::save_as_image(ft, path = "table/weather_numeric_summary_table.png")

```


\subsubsection*{2. Preprocessing and Transformation}

\begin{itemize}
    \item \textbf{Log and Differencing:} Raw cocoa prices \( P_t \) were log-transformed and differenced to obtain returns:
    \[
    r_t = \log(P_t) - \log(P_{t-1})
    \]
    ADF and KPSS tests confirmed that \( \log(P_t) \) was non-stationary while \( r_t \) was stationary.

    \item \textbf{Weather Stationarity Checks:} Weather variables were tested individually. When conflicting or non-stationary (e.g., PRCP, TAVG, TMIN), they were differenced. TMAX was retained in level form.

    \item \textbf{Peak Removal via Z-score:} To avoid distortion from extreme price spikes, a Z-score threshold of 3 was applied to \( r_t \). Observations exceeding this threshold and the following 5 days were treated as NA and interpolated.

    \item \textbf{Train-Test Split:} An 80/20 temporal split was used. All regressors were lag-aligned with the return series.
    
    \item \textbf{Missing Value Processing:} Temperature and precipitation data are partially missing for some days and were summarized using the mean across stations.
    
    \item \textbf{Merging two Datasets:} Using left join on the prices and weather datasets, and matching rows where dates are equal in both datasets with only weather data those have matching dates kept.
\end{itemize}

